---
title: "Homework Assignment 1"
author: "Ben Arancibia"
date: "September 7, 2014"
output: html_document
---
Problem Set 1
(1) Calculate the dot product of u.v
```{r}
u = matrix(c(0.5, 0.5), nrow=2, ncol=1)
u

v = matrix(c(3, -4), nrow=2, ncol=1)
v

u*v

1.5 + -2.0


```

Dot Product = -0.5

(2) What are the lengths of u and v?
Norm of a vector is ||vector||
```{r}
norm_u <- sqrt(sum(u*u))

norm_v <- sqrt(sum(v*v))

norm_u

norm_v

```

(3) What is the linear combination: 3u-2v?
Linear combination is:
```{r}
(3*u) - (2*v)
```

(4) What's the angle between u and v?
Dot-product between two vectors is proportional to their lengths and to the cosine of the angle between them. Two unit vectors, then the dot product between them is the 
cosine of the angle between them. 

Angle = Cosine(dotproduct) /||u||*||v||
```{r}
angle <- acos( sum(u*v) / ( sqrt(sum(u * u)) * sqrt(sum(v * v)) ) )

angle
```


Problem Set 2
There are two ways to solve this problem. You can use the solve function that is part of R packages or you can create your own function.

Ax = b
A*b = x
ab = x

To use the solve function you need to fill in the matrix a multiplied by b into function and column numbers.

formula you could use to calculate
A <- matrix(c(), ncol=)
b <- matrix(c(), ncol=)
x <- solve(A) %x% b

```{r}
#Example#

A <- matrix(c(1, 2, -1, 1, -1, -2, 3, 5, 4), ncol=3)  ####fill in matrix
b <- matrix (c(1, 2, 6), ncol=1)  ####fill in matrix
x <- solve (A) %*% b
x

```

