---
title: "Week 5 Assignment"
author: "Ben Arancibia"
date: "October 4, 2014"
output: pdf_document
---

#Problem Set 1
__*Question 1*__

Script to compute $A^T$$A$ and $A^T$$b$
Use the given matrix in the Problem set which is unsolvable

```{r}
A <- matrix(c(1,1,1,1,0,1,3,4), ncol=2)
b <- c(0,8,8,19)

A
b
```

Compute $A^T$$A$
```{r}
(AtA <- t(A) %*% A)
```

Compute $A^T$$b$
```{r}
(Atb <- t(A) %*% b)
```

Solve for $\hat{x}$ by doing $A^T$$A$^{-1} divided by $A^T$$b$
```{r}
(solve(AtA) %*% Atb)
```

Compute the error $e$ give the change in the variable which is now $p$
```{r}
p <- c(1, 5, 13, 17)
(Atp <- t(A) %*% p)
(x_hat <- solve(AtA) %*% Atp)
(e <- p - A %*% x_hat)
```

Find the error e=b-p
```{r}
(e = b - p)
```

Check if e is orthogonal to P and each col of A
```{r}
all.equal(e %*% p, p %*% e)
all.equal(e %*% A[,1], A[,1] %*% e)
all.equal(e %*% A[,2], A[,2] %*% e)
```

__*Question 2*__
Take the auto-mpg data extracts to an $A$ matric from the first 4 columns and $b$ vector from the fifth (*mpg*) column.
Read in the table.
__Note you will need to change your Working Directory to your local for this to work.__
```{r}
setwd('/users/bcarancibia/CUNY_IS_605')
data <- read.table("auto-mpg.data")
```

Name the columns
```{r}
names(data) <- c("displacement", "horsepower", "weight", "acceleration", "mpg")
```

Divide up the data into two matrices so that it is easier to manipulate.
```{r}
A <- as.matrix(data[,1:4])
b <- as.matrix(data[,5])
```

Use a least squares methodology and then calculate the $\hat{x}$ in order to find the best fitting equation.
```{r}
AtA <- t(A) %*% A
Atb <- t(A) %*% b
(x_hat <- solve(AtA) %*% Atb)
```

Find the error in prediction
```{r}
(e <- sqrt(sum(((A %*% x_hat) - b)^2)))
```
