---
title: "Week 7 Problem Set"
author: "Ben Arancibia"
date: "October 18, 2014"
output: html_document
---

__**Problem Set 1**__

A function to compute the expected value and standard deviation of an array of values.
```{r}
m <- array(runif(100),dim=c(10,10))
m

expected_value <- function(array){
  sum(array)/(dim(array)*dim(array))
}
expected_value(m)
mean(m)
```

Standard deviation = $Variance(x)$^$(1/2)$
```{r}
var_array <- var(m)
var_array

sd_array <- sqrt(var_array)

fixed_sd <- sd_array[!is.nan(sd_array)]

sd_m <- sqrt(fixed_sd)
sd_m
sd(m)

```

__*Part 2*__

Infinite stream of numbers coming by and need to calculate the mean and standard deviation of the numbers. I would do this by just adding to the pervious average and dividing it by n+1.

```{r}
starting_array <- array(runif(100),dim=c(10,10))
mean(starting_array)

infinite_average <- function(x,n){
  (mean(starting_array)*x)/(n+1)  
  return(infinite_average)
}

infinite_average

infinite_variance <- function(x){
  sqrt(x-infinite_average)
  return(infinite_variance)
}

infinite_variance

infinite_sd <- function(infinite_variance){
  sqrt(infinite_variance)
  return(infinite_sd)
}

infinite_sd

```


**Note**

I had a lot of issues with this homework. One of the main issues I had was I upgraded my local environment to the new OS and somehow I cannot get my TeX package to be recognized by R and print out the PDF. The HTML works fine, but I cannot get PDF. Not sure why and this is just threw me off the entire homework.

After some research found this issue: https://support.rstudio.com/hc/communities/public/questions/201688563-Problem-with-Mac-Yosemite-Beta-and-RStudio-not-finding-TeX-distribution