---
title: "Week 10 Assignment"
author: "Ben Arancibia"
date: "November 7, 2014"
output: pdf_document
---
**Playing with PageRank**

**Bullet 1**

Form the $A$ Matrix.

```{r}
A <- matrix(c(0, 0, .33, 0, 0, 0, .5, 0, .33, 0, 0, 0, .5, 0, 0, 0 ,0 ,0, 0, 0, 0, 0, 
              .5, 1, 0, 0, .33, .5, 0 ,0, 0, 0, 0, .5, .5, 0), nrow=6,ncol=6)

A


```

Introduce Decay and form the B Matrix.

```{r}
n <- 6 #chose 6 because there are 6 pages in the notes
B = 0.85*A+(0.15/n)

B
```

**Bullet 2**

Start uniform rank r and perform power iterations on B til convergence.
```{r}
r <- matrix(c(.167, .167, .167, .167, .167, .167), nrow=6, ncol=1)
r

r_solve1 = (B^10)%*%r
r_solve2 = (B^100)%*%r
r_solve3 = (B^1000)%*%r

```

**Bullet 3**
Compute the Eigendecomposition of B. Found this website to be able to do it: https://stat.ethz.ch/R-manual/R-devel/library/base/html/eigen.html

```{r}
e_decomp <- eigen(B)

e_decomp
```


**Bullet 4**
```{r}
require(igraph)
```

Use the graph package in R and its page.rank method to compute the Page Rank of the graph as given in A. Documents for igraph for future reference: http://igraph.org/r/doc/igraph.pdf

```{r}
A
net = graph.adjacency(A, mode="directed")
A_page_rank = page.rank(net, directed=TRUE)$vector


```







