---
title: "Assignment 4"
author: "Ben Arancibia"
date: "September 27, 2014"
output: html_document
---

##Problem Set 1
###Question 1 \n

We will verify that Singular Value Decomposition and eigenvalues are related. \n

Given a 3x2 martix $A$
```{r}
A <- matrix(c(1,2,3,
              -1,0,4), byrow=T, ncol=3)

A
```
\n

We will use R to compute $X$ = $AA^T$ and $Y$ = $A^TA$ and show the eigenvalues and eigenvectors of $X$ and $Y$.\n

$X$ first:
```{r}
X <- A %*% t(A)
X
```
\n

Now $Y$:
```{r}
Y <- t(A) %*% A
Y
```
\n

Eigenvalues and eigenvectors for $X$
```{r}
eigen(X)$values
eigen(X)$vectors
```
\n

Eigenvalues and Eigenvectors for $Y$
```{r}
eigen(Y)$values
eigen(Y)$vectors
```
\n

Single Value Decomposition (SVD) of $A$. Singular values vector:
```{r}
svd(A)$d
```
\n

Left Singular vector:
```{r}
svd(A)$u
```
\n

Right Singular vector:
```{r}
svd(A)$v
```
\n


$X$ and $Y$ eigens are equal
```{r}
all.equal(eigen(X)$values, eigen(Y)$values[c(1,2)])
```
\n

Showing that the SVD vectors are equal:
```{r}
all.equal(svd(A)$v[,1], (eigen(Y)$vectors[,1]))
all.equal(svd(A)$v[,2], (eigen(Y)$vectors[,2]))
all.equal(svd(A)$u[,1], (eigen(X)$vectors[,1]))
all.equal(svd(A)$u[,2], (eigen(X)$vectors[,2]))
```
\n
For some reason they are not equal....unless you reverse the signs. Why? \n
# Need to ask professor about this \n

```{r}
all.equal(svd(A)$v[,1], (-1 * eigen(Y)$vectors[,1]))
all.equal(svd(A)$u[,1], (-1 * eigen(X)$vectors[,1]))
```

##Problem Set 2
###Question 1

Writing a function computer inverse of well-conditioned full-rank square matrix using co-factors.\n
Use the following signature:
\n
B = myinverse(A)
\n
$A$ is a matrix and $B$ is its inverse and $A$x$B$ = $I$\n
# for this to work, matrix needs to be invertible. did not build in test for invertible matrix, would do that in the future assignment \n

function:
```{r}
myinverse <- function(m){
  x <- matrix(ncol=ncol(m), nrow=nrow(m))
  for (i in 1:nrow(m)){
    for (j in 1:ncol(m)){
      x[i,j] <- (-1)^(i+j) * det(m[-i,-j])
    }
  }
  return(t(x)/det(m))
}

```
\n

test invertible matrix (got invertible matrix from http://www.purplemath.com/modules/mtrxinvr2.htm):
```{r}
A <- matrix(c(1,2,3,
              0,1,4,
              5,6,0), nrow=3, byrow=T)
B <- myinverse(A)
B
```

test $A$$B$=$I$:
```{r}
I <- diag(3)
all.equal(A %*% B, diag(3))
all.equal(B, solve(A))
```

