---
title: "Assignment 3"
author: "Ben Arancibia"
date: "September 20, 2014"
output: html_document
---
```{r, include=FALSE}
library(pracma)
library(Matrix)#matrix
```

#Problem Set 1
###Question 1
Rank of Matrix A is 4 since there are 4 linearly indepedent rows. None of the rows can be written as a linear combination(quick explanation of linear independence for my reference: http://en.wikipedia.org/wiki/Linear_independence).

###Question 2
Given an mxn matrix where m > n then the maximum rank is n, you cannot have more than n linearly independent rows since m > n. Assuming matrix is a non-zero, the minimum rank is 1.

###Question 3
The rank of matrix B is 1.
```{r}
##### Question 3 #####
m2 <- matrix(c(1,2,1,
               3,6,3,
               2,4,2), byrow=T, nrow=3)
rankMatrix(m2)[1]
rref(m2)
```

#Problem set 2
###Question 1
```{r}
q1 <- matrix(c(1,2,3,0,4,5,0,0,6), byrow=T, nrow=3)
q1

```

```{r}
# matrix is in upper triangular form, eigenvalues are simply the diagonals: 6, 4, 1
eigen(q1)$values
eigen(q1)$vectors
```

Manual Computation \n

$Av = \lambda v$ \n

$(A - \lambda I)v = 0$ \n

Matrix determinant $A - \lambda I=0$ \n

Begin computation
$$det \left( \begin{pmatrix}1&2&3\\0&4&5 \\ 0&0&6 \end{pmatrix} - \begin{pmatrix}\lambda&0&0\\0&\lambda&0\\0&0&\lambda\end{pmatrix}\right)=det\begin{pmatrix}1 - \lambda&2&3\\0&4 - \lambda&5 \\ 0&0&6 - \lambda \end{pmatrix}$$
$$ (1 - \lambda)[(4-\lambda)(6-\lambda)-(5)(0)]-2[(0)(6-\lambda)-(5)(0)]-3[(0)(0) - (4 - \lambda)(0)]=0 $$ Which gives us $6,4,1$ \n

Find $AX=0$ \n

Solve for eigenvalue 6:
$$\begin{pmatrix}1 - 6&2&3\\0&4-6&5 \\ 0&0&6-6 \end{pmatrix}=\begin{pmatrix}-5&2&3\\0&-2&5 \\ 0&0&0 \end{pmatrix}$$ \n

Solve eigenvector: \n

$$\begin{pmatrix}-5&2&3\\0&-2&5 \\ 0&0&0 \end{pmatrix}\begin{pmatrix}v_1\\v_2\\v_3\end{pmatrix}=\begin{pmatrix}0\\0\\0\end{pmatrix}$$
$$-5v_1 + 2v_2 +3v_3=0$$
$$0v_1 - 2v_2 + 5v_3 = 0$$
$$0v_1 + 0v_2 + 0v_3 = 0$$ \n

$$2v_2 = 5v_3 \sim v_2 =\frac{5}{2}v_3$$\n

$$ -5v_1 + 2(\frac{5}{2}v_3) +3v_3 = -5v_1+8v_3 \sim v_1 = \frac{8}{5} v_3 $$ \n

Eigenvector of eigenvalue 6: $$\begin{pmatrix}\frac{8}{5}&\frac{5}{2}&1\end{pmatrix}$$ \n

Solve for eigenvalue 4 \n
$$\begin{pmatrix}-3&2&3\\0&0&5\\0&0&2\end{pmatrix}$$ Eigenvector of eigenvalue 4: $$\begin{pmatrix}\frac{2}{3} & 1 & 0\end{pmatrix}$$ \n

Solve for eigenvalue 1:\n

$$\begin{pmatrix}0&2&3\\0&3&5\\0&0&5\end{pmatrix}$$ Eigenvector of eigenvalue 1: $$\begin{pmatrix}1&0&0\end{pmatrix}$$
