---
title: "Assignment 13"
author: "Ben Arancibia"
date: "November 23, 2014"
output: pdf_document
---

__**Question 1**__

Write a program to compute the derivative of $f(x) = x^3 + 2x^2$. The function should take in a value of x and return back an approximation to the derivate of f(x) evaluated at the value.

```{r}
invisible(library(mosaic))

f1 <- D(x^3 + 2*x^2 ~ x)

#as example can solve for any value
f1(x=2)

```

__**Question 2**__

Now write a program to compute the area under the curve for the function $2X^2 + 4x$ in the range $x$ = [1,3].

```{r}
library(zoo)

f2 <- function(x) (2*x^2 + 4*x)
xs <- seq(1, 3, by = 0.000006)
plot(xs, f2(xs), type= "l")


x <- 1:3
y <- 2*x^2 + 4*x
id <- order(x)
AUC <- sum(diff(x[id])*rollmean(y[id],2))

AUC

```


__**Question 3**__

**Integration by parts to solve for $$\int sin(x)cos(x)\, dx$$**

$$=1/2 sin2x$$

$$\int 1/2 sin2x\, dx $$

final answer: $$-1/2 cos^2(x) + C$$


**Integration by parts to solve for $$\int x^2e^x\, dx$$**

$$ u dv = uv - integral(v du)$$

u = x^2 
du = 2x dx
dv = e^x dx
v = e^x

$$x^2 * e^x - integral(2x * e^x dx)$$

We have to do it again

u = x
du = dx
dv = e^x dx
v = e^x

$$x^2 * e^x - 2[x*e^x - integral(e^x dx)]$$

Simplify:

$$x^2 * e^x - 2x*e^x + 2*integral(e^x dx) $$


$$x^2*e^x - 2x*e^x + 2*e^x + C$$

Factor out: e^x


Final Answer: $$e^x [x^2 - 2x + 2] + C$$


**What is the derivative of $$xcos(x)$$**

Use the product rule.

$$(f*g) = f'*g + g'*f$$

f' = 1
g'=-sin(x)

$$1*cos(x) - xsin(x)$$

Final Answer: $$cos(x)-xsin(x)$$


**What is the derivative of e^x^4**

e^x^4 * d/dx (x^4)

Final Answer: 4x^3 * e^x^4








