---
title: "Assignment 9"
author: "Ben Arancibia"
date: "November 1, 2014"
output: pdf_document
---

__**Problem Set 1**__

1) Function that will produce a sample of random variable that is distrubuted as follows: 
$$f(x) = \begin{cases}
  x & \text{if } 0 \leq x \leq 1 \\
  2-x & \text{if } 1 \leq x \leq 2
  \end{cases}
$$

```{r}
function_one <- function(x) { 
  if (x >= 0 && x <= 1) {
    return(x)
  } else if (x > 1 && x <= 2) {
    return(2-x)
  }
}

```

2) Function that will produce a sample of random variable that is distrubuted as follows: 
$$f(x) = \begin{cases}
  1-x & \text{if } 0 \leq x \leq 1 \\
  x-1 & \text{if } 1 \leq x \leq 2
  \end{cases}
$$

```{r}
function_two <- function(x) { 
  if (x >= 0 && x <= 1) {
    return(1-x)
  } else if (x > 1 && x <= 2) {
    return(x-1)
  }
}
```

3) 1000 samples using each pdf. Save one histogram for each PDF.

```{r}
sample_one <- sample(
  seq(0,2, by=0.01), 1000, replace=T, 
  prob=sapply(
    seq(0,2,by=0.01), 
    function(x) function_one(x)
  )
)

hist(sample_one, 30, freq=F)

sample_two <- sample(
  seq(0,2, by=0.01), 1000, replace=T, 
  prob=sapply(
    seq(0,2,by=0.01), 
    function(x) function_two(x)
  )
)

hist(sample_two, 30, freq=F)

```

4) Size n as a parameter and the PDF as the second parameter, perform 1000 iterations. Plot a histogram.

```{r}
problem_four <- function(n, pdf) {
  nx <- sample(seq(0,2, by=0.01), n, replace=T,
               prob = sapply(
                 seq(0,2,by=0.01),
                 function(x) {
                   pdf(x)}))
  hist(nx, 30, freq=F)
  print(paste("The sample mean for", n, "samples from the pdf is", mean(nx)))
}

problem_four(10000, function_one)

problem_four(10000, function_two)

```

5) Use the PDFs with sample sizes of 10 and 20.

PDF 1
```{r}
problem_four (1000, function_one)
problem_four (20, function_one)
problem_four (10, function_one)

```

PDF 2
```{r}
problem_four (1000, function_two)
problem_four (20, function_two)
problem_four (10, function_two)

```