---
title: "Assignment 8"
author: "Ben Arancibia"
date: "October 25, 2014"
output: pdf_document
---

__**Problem Set 1**__

Probability of $(A|B)$ is equal to $PR(B|A)$*$PR(A)$ / $PR(B)$.

So $PR(B)$ = ((.9x10) + (.5x10))/20

14/20 = $0.7$

$PR(B|A)$ = 0.5

$PR(A)$ = 0.5

$PR(A|B)$ = (0.5 x 0.5)/(0.7)

(0.25)/(0.7)

__final result:__
$0.3571$

__**Problem Set 2**__
What happens to the probability of Difficulty of Course when you present the evidence that the received recommendation letter was good?


```{r}

##FUTURE REFERENCE for gRain: http://www.bioconductor.org/packages/release/bioc/html/RBGL.html##

library(gRain)
yes_no <- c("yes", "no")
high_low <- c("high", "low")

#percents from variables 
diff <- cptable(~difficulty, values=c(70, 30), levels=yes_no)
intell <- cptable(~intelligence, values=c(20, 80), levels=high_low)

grade_intell_diff <- cptable(~grade|intelligence:difficulty, 
                             values=c(90, 10, 99, 1, 20, 80, 40, 60), levels=high_low)

sat_intell <- cptable(~sat|intelligence, values=c(80, 20, 10, 90), levels=high_low)
letter_grade <- cptable(~letter|grade, values=c(95, 5, 10, 90), levels=high_low)

p_var <- compileCPT(list(diff, intell, grade_intell_diff, sat_intell, letter_grade))

#create bayesian network
bay_network <- grain(p_var)

p_var$difficulty

```

p_var$difficult seems to have worked, which is just probability.

A course seems to be more difficult given that a high recommendation letter was given. (0.84 > 0.7)
```{r}
bay_network_LH <- setFinding(bay_network, nodes="letter", states="high")
#query grain time
querygrain(bay_network_LH, nodes=c("letter", "difficulty"))$difficulty
```

Seems that a course was difficult given a high SAT Score and a high recommendation, than just a high recommendation alone. (0.86 > 0.84)

```{r}
bay_network_LHSH <- setFinding(bay_network, nodes=c("sat", "letter"), states=c("high", "high"))
#query grain time
querygrain(bay_network_LHSH, nodes=c("letter", "sat", "difficulty"), type="marginal")$difficulty
```

