---
title: "Week 11 Assignment"
author: "Ben Arancibia"
date: "November 15, 2014"
output: pdf_document
---

Using the lm function, perform regression analysis and measure independent variables on two datasets.

__**First Data Set**__
The first data set is heart rate.
First create the data set.
```{r}
x = c(18,23,25,35,65,54,34,56,72,19,23,42,18,39,37)
y = c(202,186,187,180,156,169,174,172,153,199,193,174,198,183,178) 
```

Plot x and y with regression line and basic values of regression analysis
```{r}
plot(x,y)
lm_age = lm(y ~ x)
lm_age
abline(lm_age)
summary(lm_age)

```

Find that the resulting equation is more like this:

$MaxHR = -0.7977 + 210.0485$ 

As you can see in summary(lm_age) you have a hypothesis test calculated by R.
```{r}
summary(lm_age)
```
It does not look like it is significant. 

```{r}
es = residuals(lm_age)
b1 = (coef(lm_age))[['x']]
s = sqrt(sum(es^2) / (13)) #plugging in n-2 instead of 13 gives me an error
SE = s/sqrt(sum((x-mean(x))^2))
t = (b1 - (-1))/SE
pt(t, 13, lower.tail=FALSE)

```

Not significiant.

__**Auto Data Set**__

Perform a linear regression analysis using mpg as the dependent variable and the other 4 as independent variables.
First import the data.

```{r}
data <- read.table("/Users/bcarancibia/CUNY_IS_605/assign11/auto-mpg.data")
names(data) <- c("displacement", "horsepower", "weight", "acceleration", "mpg")
```

Based on the first take a random 40 points from the data set.
```{r}
sub_new <- data[1:40,]
```

Plot mpg vs the four other variables for the subset.
```{r}
pairs(sub_new)
auto_sub <- lm(sub_new$mpg ~ sub_new$displacement + sub_new$horsepower+ sub_new$weight +
                 sub_new$acceleration, data=sub_new)

auto_sub

layout(matrix(c(1,2,3,4),2,2))
plot(auto_sub)
summary(auto_sub)
```

*The final regression fit (40 data points) looks to be:*

$mpg = 41.628 - .0197displacement - .00733horsepower - .00371weight - .365acceleration$

Weight has a sifnifcant impact on mpg (if you look at at summary auto_sub)

*The corresponding significance levels are:*

displacement: 0.014606

horsepower: 0.607200

weight: .000207

acceleration: .045472

*Standard errors for each of the coefficients:*

displacement: 0.00767703

horsepower: 0.0141243

weight: 0.0008949

acceleration: 0.1757330



Plot mpg vs the four other variables.
```{r}

pairs(data)

auto <- lm(data$mpg ~ data$displacement + data$horsepower 
           + data$weight+ data$acceleration, data=data)

auto

layout(matrix(c(1,2,3,4),2,2))
plot(auto)

summary(auto)

```

*The final regression fit (all data) looks to be:*

$mpg = 45.25 - 0.006001displacement - 0.0436007horsepower - 0.005281weight - 0.023148accerleration

Again weight seems to have the most signifcant impact on mpg, but horsepower also is significant.

*The corresponding significance levels are:*

displacement:0.37166

horsepower: 0.00885

weight: 2.3e-10

acceleration: 0.85388


*Standard errors for each of the coefficients:*

displacement: 0.0067093

horsepower: 0.0165735

weight: 0.0008109

acceleration: 0.1256012

__Measure the the 95% confidence intervals__

```{r}

confint(auto_sub, level=0.95)


confint(auto, level=0.95)
```
